---
title: "unittesting"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Unit Testing


Load all the functions from the user interface
```{r}
#setwd("~/fit3164/webapp/")
#source("userInputs_ui.R")
library(testthat)
```

Unit testing for calculateBMI
This will test the accuracy of the calculateBMI function, by comparing the outputs of the function to expected BMI
inputs: calculateBMI() function, 
        these weight/height combinations: (118, 167), (124.6, 217), (49.88472,153.7757)
expected output: No output if all the tests are correct

```{r}

test_that("BMI of user",
          { expect_equal(calculateBMI(118, 167), 42.310588404)
            expect_equal(calculateBMI(124.6, 217), 26.4605321837)
            expect_equal(calculateBMI(49.88472,153.7757),21.0956112462)

          })
```

Unit testing for isObese
This will test the accuracy of the isObese function, by comparing the outputs of the function to expected output
inputs: isObese() function, 
        these BMIs: 42, 30, 29.5, 26
expected output: No output if all the tests are correct
```{r}
test_that("Obesity of user",
          { expect_equal(isObese(42), "Y")
            expect_equal(isObese(30), "Y")
            expect_equal(isObese(29.5),"N")
            expect_equal(isObese(26),"N")

          })
```


Unit testing for getGender
This will test the accuracy of the getGender function, by comparing the outputs of the function to expected gender
inputs: getGender() function, 
        "Female"
        "Male"
expected output: No output if all the tests are correct
```{r}
test_that("Gender of user",
          { expect_equal(getGender("Female"), "Fmale")
            expect_equal(getGender("Male"), "Male")

          })
```

Unit testing for changeValues
This will test the accuracy of the changeValues function, by comparing the outputs of the function to expected output
inputs: changeValues() function, 
        "Yes"
        "No"
expected output: No output if all the tests are correct
```{r}
test_that("Changing values",
          { expect_equal(changeValues("Yes"), 1)
            expect_equal(changeValues("No"), 0)

          })
```

Unit testing for getBBB
This will test the accuracy of the getBBB function, by comparing the outputs of the function to expected output
inputs: getBBB() function, 
        "Left"
        "Right"
        "None"
expected output: No output if all the tests are correct
```{r}
test_that("BBB",
          { expect_equal(getBBB("Left"), "LBBB")
            expect_equal(getBBB("None"), "N")
            expect_equal(getBBB("Right"), "RBBB")

          })
```

Unit testing for getVHD
This will test the accuracy of the getVHD function, by comparing the outputs of the function to expected output
inputs: getVHD() function 
        "Mild"
        "Moderate"
        "None"
        "Severe"
expected output: No output if all the tests are correct
```{r}
test_that("VHD",
          { expect_equal(getVHD("Mild"), "mild")
            expect_equal(getVHD("None"), "N")
            expect_equal(getVHD("Moderate"), "Moderate")
            expect_equal(getVHD("Severe"), "Severe")

          })
```