---
title: "unusedcode"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The following code is not used for our project.

# Data processing:

Hungary data set
```{r}

# this is the 76 variable set

#this line works
h<-readtext("hungarian.data")
h<- h[-1]


#split rows

h<-h %>% 
  separate_rows(text, sep = "name")
# split columns

# remove \n from row2:

h<-h %>%
  mutate(text = str_replace(text, "\n",""))


h<-h %>% separate(text, c("id","ccf", "age" ,"sex", "painloc" ,"painexer" ,"relrest" ,"pncaden" ,"cp", "trestbps","htn" , "chol","smoke" ,"cigs" , "years" ,"fbs", "dm" ,"famhist" ,"restecg", "ekgmo" ,"ekgday" , "ekgyr" ,"dig" ,"prop" ,"nitr" ,"pro" ,"diuretic" ,"proto" ,"thaldur" , "thaltime" ,"mets" ,"thalach","thalrest" ,"tpeaksbps" ,"tpeaksbpd" , "dummy" ,"trestbpd" ,"exang","xhypo" , "oldpeak", "slope", "r1dv5" ,"r1dv5e" ,"ca", "restckm" ,"exerckm" ,"restef" ,"restwm" ,"exeref", "exerwm"  ,"thal", "thalsev" , "thalpul" ,"earlobe" ,"cmo" ,"cday" ,"cyr" ,"num", "lmt" ,"ladprox" ,"laddist" , "diag" ,"cxmain" , "ramus" ,"om1" ,"om2" ,"rcaprox" ,"rcadist" ,"lvx1" ,"lvx2" ,"lvx3" ,"lvx4" ,"lvf" ,"cathef" ,"junk"))

# we are not using this anymore so we will not save it
#write.csv(h,'hungary_large.csv')

hungarian<-h
```

Cleveland set
```{r}
# large set

# this is the 76 variable set

#this line works
c<-readtext("cleveland.data")
c<- c[-1]


#split rows

c<-c %>% 
  separate_rows(text, sep = "name")
# split columns

# remove \n from row2:

c<-c %>%
  mutate(text = str_replace(text, "\n",""))


c<-c %>% separate(text, c("id","ccf", "age" ,"sex", "painloc" ,"painexer" ,"relrest" ,"pncaden" ,"cp", "trestbps","htn" , "chol","smoke" ,"cigs" , "years" ,"fbs", "dm" ,"famhist" ,"restecg", "ekgmo" ,"ekgday" , "ekgyr" ,"dig" ,"prop" ,"nitr" ,"pro" ,"diuretic" ,"proto" ,"thaldur" , "thaltime" ,"mets" ,"thalach","thalrest" ,"tpeaksbps" ,"tpeaksbpd" , "dummy" ,"trestbpd" ,"exang","xhypo" , "oldpeak", "slope", "r1dv5" ,"r1dv5e" ,"ca", "restckm" ,"exerckm" ,"restef" ,"restwm" ,"exeref", "exerwm"  ,"thal", "thalsev" , "thalpul" ,"earlobe" ,"cmo" ,"cday" ,"cyr" ,"num", "lmt" ,"ladprox" ,"laddist" , "diag" ,"cxmain" , "ramus" ,"om1" ,"om2" ,"rcaprox" ,"rcadist" ,"lvx1" ,"lvx2" ,"lvx3" ,"lvx4" ,"lvf" ,"cathef" ,"junk"))

#write.csv(c,'cleveland_large.csv')

cleveland<-c
```

switzerland set
```{r}
# large set

# this is the 76 variable set

#this line works
c<-readtext("switzerland.data")
c<- c[-1]


#split rows

c<-c %>% 
  separate_rows(text, sep = "name")
# split columns

# remove \n from row2:

c<-c %>%
  mutate(text = str_replace(text, "\n",""))


c<-c %>% separate(text, c("id","ccf", "age" ,"sex", "painloc" ,"painexer" ,"relrest" ,"pncaden" ,"cp", "trestbps","htn" , "chol","smoke" ,"cigs" , "years" ,"fbs", "dm" ,"famhist" ,"restecg", "ekgmo" ,"ekgday" , "ekgyr" ,"dig" ,"prop" ,"nitr" ,"pro" ,"diuretic" ,"proto" ,"thaldur" , "thaltime" ,"mets" ,"thalach","thalrest" ,"tpeaksbps" ,"tpeaksbpd" , "dummy" ,"trestbpd" ,"exang","xhypo" , "oldpeak", "slope", "r1dv5" ,"r1dv5e" ,"ca", "restckm" ,"exerckm" ,"restef" ,"restwm" ,"exeref", "exerwm"  ,"thal", "thalsev" , "thalpul" ,"earlobe" ,"cmo" ,"cday" ,"cyr" ,"num", "lmt" ,"ladprox" ,"laddist" , "diag" ,"cxmain" , "ramus" ,"om1" ,"om2" ,"rcaprox" ,"rcadist" ,"lvx1" ,"lvx2" ,"lvx3" ,"lvx4" ,"lvf" ,"cathef" ,"junk"))

#write.csv(c,'switzerland_large.csv')

switzerland<-c
```

long beach set
```{r}
# large set

# this is the  76 variable set

#this line works
c<-readtext("long-beach-va.data")
c<- c[-1]


#split rows

c<-c %>% 
  separate_rows(text, sep = "name")
# split columns

# remove \n from row2:

c<-c %>%
  mutate(text = str_replace(text, "\n",""))


c<-c %>% separate(text, c("id","ccf", "age" ,"sex", "painloc" ,"painexer" ,"relrest" ,"pncaden" ,"cp", "trestbps","htn" , "chol","smoke" ,"cigs" , "years" ,"fbs", "dm" ,"famhist" ,"restecg", "ekgmo" ,"ekgday" , "ekgyr" ,"dig" ,"prop" ,"nitr" ,"pro" ,"diuretic" ,"proto" ,"thaldur" , "thaltime" ,"mets" ,"thalach","thalrest" ,"tpeaksbps" ,"tpeaksbpd" , "dummy" ,"trestbpd" ,"exang","xhypo" , "oldpeak", "slope", "r1dv5" ,"r1dv5e" ,"ca", "restckm" ,"exerckm" ,"restef" ,"restwm" ,"exeref", "exerwm"  ,"thal", "thalsev" , "thalpul" ,"earlobe" ,"cmo" ,"cday" ,"cyr" ,"num", "lmt" ,"ladprox" ,"laddist" , "diag" ,"cxmain" , "ramus" ,"om1" ,"om2" ,"rcaprox" ,"rcadist" ,"lvx1" ,"lvx2" ,"lvx3" ,"lvx4" ,"lvf" ,"cathef" ,"junk"))

write.csv(c,'longbeach_large.csv')

longbeach<-c
```

Z sani set                4 data sets
Typical.Chest.Pain        derive from cp value 1
Atypical                  derive from cp value 2
Age                       age
HTN                       hypertension- blood pressure above 130/80 - 140/90mmgg(mayo clinic definition)
                          I'm making it trestbpd>80 AND trestbps>130 = YES
BP                        blood pressure- most likely trestbps
Current.Smoker            smoke 1 = yes; 0 = no 
FBS                       fbs (fasting blood sugar > 120 mg/dl)  (1 = true; 0 = false)


```{r}
# Add all the extra data together

other <- rbind(cleveland, switzerland)
other<- rbind(other, longbeach)
other<- rbind(other, hungarian)


# delete repeated rows
other<-distinct(other)

# formatting
other$id<-NULL
other$ccf<-NULL
other$painloc<-NULL
other$dummy<-NULL
other$lvx1<-NULL
other$lvx2<-NULL
other$lvx3<-NULL
other$lvx4<-NULL
other$cathef<-NULL
other$junk<-NULL

# change age to numeric, then delete incorrect rows
other$age<- as.numeric(other$age)

# change sex to factor form
other$sex[other$sex == 1]<- "Male"
other$sex[other$sex == 0]<- "Fmale"
other$sex<-as.factor(other$sex)

# change painexer
# 9 is the NA value
other$painexer[other$painexer == 9]<- NA
# 1=Y 0=N
other$painexer[other$painexer == 1]<- "Y"
other$painexer[other$painexer == 0]<- "N"

# rename some columns
#HTN  =hypertension- blood pressure above 130/80 - 140/90mmgg(mayo clinic definition)

other <- other %>% rename(Exertional.CP= painexer, Age=age, Sex=sex, HTN=htn)

#Typical.Chest.Pain -derive from cp=value 1
other$Typical.Chest.Pain <- 0
other$Typical.Chest.Pain[other$cp == 1]<- 1

# Atypical                  derive from cp value 2
other$Atypical<- 0
other$Atypical[other$cp == 2]<- 1

#BP                        blood pressure- most likely trestbps
other <- other %>% rename(BP=trestbps)
#Current.Smoker            smoke 1 = yes; 0 = no 
other <- other %>% rename(Current.Smoker=smoke)

# Cath  Cad or Normal        =          num Value 0: < 50% diameter narrowing, Value 1: > 50% diameter narrowing
# most of the num values are not 1 or 0? I am choosing to delete those rows
other$Cath<- 0
# if not 1 or 0 set to NA
other$Cath[other$num>1]<-NA
other$Cath[other$num==0]<- "Normal"
other$Cath[other$num==1]<- "Cath"

# select the columns of the other data sets to use
suplementary<- other %>%select(Sex, Age, Exertional.CP, HTN, Atypical, Typical.Chest.Pain, BP, Current.Smoker, fbs, Cath)

# change the z sani data set to the form in the comment above
# fbs (fasting blood sugar > 120 mg/dl)  (1 = true; 0 = false)
z2<-z
z2$fbs<-0
z2$fbs[z2$FBS>120]<-1


# select the columns
z3 <- select(z2, Sex, Age, Exertional.CP, HTN, Atypical, Typical.Chest.Pain, BP, Current.Smoker, fbs, Cath)
# join the two sets
# s=factor age=num ext=factor htn=num (change to fact) atyp=factor typ=num (change fac) bp=num current=factor fbs=num (fac) Cath=factor
z3$HTN<-as.factor(z3$HTN)
z3$Typical.Chest.Pain<- as.factor(z3$Typical.Chest.Pain)
z3$fbs<- as.factor(z3$fbs)

# supplementary sex remove invalid
 

combined<- rbind(z3, suplementary)
```

We will not use these data sets as there are too many missing values and some important variables are not there

# Additional unused code relating to UI

#Show variable summary on Data Overview page
getVariableSummary <- function(var){
  varName = variables[var]
  value = eval(parse(text = varName))
  # CHANGE FOR THOSE WITH NO MIN/MAX ELIGIBLE VALUES
  variable_summary = data.frame("Min"=round(min(value), digits=2), "Max"=round(max(value), digits=2), "Mean"=round(mean(value), digits=2), "Median"=round(median(value), digits=2))
  return(variable_summary)
}

#Static violin model showing gender distribution/HD
fig <- data %>%
  plot_ly(type = 'violin') 
fig <- fig %>%
  add_trace(
    x = ~Cath[data$Sex == 'Male'],
    y = ~Age[data$Sex == 'Male'],
    ylim(0,1),
    legendgroup = 'Yes',
    scalegroup = 'Yes',
    name = 'Male',
    side = 'negative',
    box = list(
      visible = T
    ),
    meanline = list(
      visible = T
    ),
    color = I("blue")
  ) 
fig <- fig %>%
  add_trace(
    x = ~Cath[data$Sex == 'Fmale'],
    y = ~Age[data$Sex == 'Fmale'],
    ylim(0,1),
    legendgroup = 'No',
    scalegroup = 'No',
    name = 'Female',
    side = 'positive',
    box = list(
      visible = T
    ),
    meanline = list(
      visible = T
    ),
    color = I("green")
  ) 

fig <- fig %>%
  layout(
    xaxis = list(
      title = ""  
    ),
    yaxis = list(
      title = "",
      zeroline = F
    ),
    violingap = 0,
    violingroupgap = 0,
    violinmode = 'overlay'
  )

#### original scatterplot of continuous variables ####
za_cont <- data.frame(Age, Weight, Length, BMI, BP, PR, CR, TG, LDL,
                      HDL, BUN, ESR, HB, K, Na, WBC, Lymph)

pearson_cor_func <- function(var_1, var_2) {
  #correlate
  cor_result <- cor.test(var_1, var_2, method = 'pearson')
  #make an interactive plot
  fig <- plot_ly(data = za_cont,
                 x = ~var_1,
                 y = ~var_2,
                 type = 'scatter',
                 text = ~paste('Variable 1:', var_1, '<br>Variable 2:', var_2), 
                 size = ~var_1)
  #return results
  result_list <- list('result' = cor_result,'int_plot' = fig)
  return(result_list)
}

pearson_cor_func(Age, Weight)

ggscatter(za, x = "Weight", y = "BMI", add = 'reg.line', 
                       conf.int = TRUE, cor.coef = TRUE, cor.methods = 'pearson',
                       xlab = 'Weight', ylab = 'BMI')

#### original bar plot of binary variables, split between healthy and heart disease ####
library(plotly)
#library(dpylr)
attach(za)

yes_cath <- filter(za, Cath == 'Cad')
no_cath <- filter(za, Cath == 'Normal')

yes_small <- data.frame(yes_cath$DM,yes_cath$Current.Smoker,yes_cath$EX.Smoker,yes_cath$FH,yes_cath$Obesity,yes_cath$CRF,yes_cath$CVA,
                       yes_cath$Airway.disease,yes_cath$Thyroid.Disease,yes_cath$CHF,yes_cath$DLP,yes_cath$Edema,yes_cath$Weak.Peripheral.Pulse,
                      yes_cath$Lung.rales,yes_cath$Systolic.Murmur,yes_cath$Diastolic.Murmur,yes_cath$Typical.Chest.Pain,yes_cath$Dyspnea,
                       yes_cath$Atypical,yes_cath$Nonanginal,yes_cath$Exertional.CP,yes_cath$LowTH.Ang,yes_cath$Q.Wave,yes_cath$St.Elevation,
                       yes_cath$St.Depression,yes_cath$Tinversion,yes_cath$LVH,yes_cath$Poor.R.Progression,yes_cath$BBB)
                        
no_small <- data.frame(no_cath$DM,no_cath$Current.Smoker,no_cath$EX.Smoker,no_cath$FH,no_cath$Obesity,no_cath$CRF,no_cath$CVA,
                       no_cath$Airway.disease,no_cath$Thyroid.Disease,no_cath$CHF,no_cath$DLP,no_cath$Edema,no_cath$Weak.Peripheral.Pulse,
                       no_cath$Lung.rales,no_cath$Systolic.Murmur,no_cath$Diastolic.Murmur,no_cath$Typical.Chest.Pain,no_cath$Dyspnea,
                       no_cath$Atypical,no_cath$Nonanginal,no_cath$Exertional.CP,no_cath$LowTH.Ang,no_cath$Q.Wave,no_cath$St.Elevation,
                       no_cath$St.Depression,no_cath$Tinversion,no_cath$LVH,no_cath$Poor.R.Progression,no_cath$BBB)

new <- merge(data.frame(yes_small, row.names = NULL), data.frame(no_small, row.names = NULL), by = 0, all = TRUE)[-1]

Variables <- c('DM','Current.Smoker','EX.Smoker','FH','Obesity','CRF','CVA',
          'Airway.disease','Thyroid.Disease','CHF','DLP','Edema','Weak.Peripheral.Pulse',
          'Lung.rales','Systolic.Murmur','Diastolic.Murmur','Typical.Chest.Pain','Dyspnea',
          'Atypical','Nonanginal','Exertional.CP','LowTH.Ang','Q.Wave','St.Elevation',
          'St.Depression','Tinversion','LVH','Poor.R.Progression','BBB')

yes_vars <- c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28,29)
no_vars <- c(30,32,33,34,35,36,37,38,39,40,41,42,43,44,45,46,47,48,49,50,51,52,53,54,55,56,57,58,59)

plot_data <- data.frame(vars, yes_vars, no_vars)
fig <- plot_ly(plot_data, x = ~Variables, y = ~yes_vars, type = 'bar', name = 'Heart Disease')
fig <- fig %>% add_trace(y = ~no_vars, name = 'Healthy')
fig <- fig %>% layout(yaxis = list(title = 'Total Number'), barmode = 'group')

fig

#### testing regression + interaction models for some initial exploration ####
#regression models
library(ggplot2)
library(car)
library(QuantPsyc)
library(lmSupport)
library(psych)
library(haven)
library(rockchalk)

#change cath to factor
z$Cath<-as.factor(z$Cath)

####simple model####
#lm(dependent variable ~ independent variable)
simple_reg <- lm(z$BP~z$Typical.Chest.Pain)

par(mfrow = c(2,2))
plot(simple_reg)

dwt(simple_reg)

summary(simple_reg, confint = TRUE)

#plot with regression line using coefficients
par(mfrow = c(1,1))
plot(z$Typical.Chest.Pain, z$BP)
abline(lm(z$BP~z$Typical.Chest.Pain))

####interaction model####
z$Cath <- to_factor(z$Cath)
z$Cath <-as.numeric(z$Cath)

interact_reg <- lm(Cath ~ BP*Typical.Chest.Pain, data = z)

par(mfrow = c(2,2))
plot(interact_reg)

dwt(interact_reg)
vif(interact_reg)

summary(interact_reg, confint = TRUE)
lm.beta(interact_reg)
lm.sumSquares(interact_reg)

#plot 
par(mfrow = c(1,1))

#library(sjPlot)
#library(sjmisc)
#theme_set(theme_sjplot())
#plot_model(interact_reg, type = 'pred')

plotSlopes(interact_reg, plotx = "BP", modx = "Typical.Chest.Pain", modxVals = "std.dev", main = 'simple slopes plot')

#interaction.plot(z$BP, z$Typical.Chest.Pain, z$Cath)
